### 2022-01-28 chapter2/chapter.tex
same draft

%!TEX root = ../thesis.tex
\chapter{Approximating fluid queues with the discontinuous Galerkin method \label{ch:galerkin}} 
% PREFACE WITH CONTRIBUTIONS 
% TUTORIAL: HOW TO -- AS IN PEPER
% PROBLEMS AND SOLUTIONS 
% 	NON-NEGATIVITY AND OSCILLATIONS 
%	FILTERING AND LIMITING 
%	NOT A SOLUTION IN OUR CONTEXT & MOTIVATION FOR A NEW METHOD -- THE QBD-RAP

>>>>

%!TEX root = ../thesis.tex
\chapter{Approximating fluid queues with the discontinuous Galerkin method \label{ch:galerkin}} 
\begin{center}
\begin{minipage}{0.8\textwidth}
    \textit{Sections~\ref{sec:DG} and \ref{sec: approx r} have been taken from Sections~4 and 5 of \cite{blnos2022} with only minor changes, such as notations, so that this chapter is consistent with the rest of the thesis. I am a co-author of the paper \cite{blnos2022}. The conceptualisation of \cite{blnos2022} was originally by Vikram Sunkara, Nigel Bean and Giang Nguyen, and the original coding was done by Vikram Sunkara. I made significant contributions to Section~3 of the paper, expressing the operator-theoretic expressions to use the same partition as the approximation scheme. I contributed Sections~4.4 and 5.1. I extended the numerical experiments in Section~6 to higher orders and made all the plots in Section~6. Appendix~A is also my original work. I did a significant proportion of the writing of the manuscript and addressed the reviewers comments and also developed code for the numerical experiments.
    }
\end{minipage}
\end{center}
% The DG method for B 
    % direct copy paste from the paper
% Problems with DG, non-neg/oscillations
    % solutions: filtering and limiting
% Approximating R: projection, interpolation, cell averages (eluding to QBD-RAP and unif method)
% Approximating Psi (and the stationary distribution, but this isnt really necessary)
    % constructing D
    % solving the matrix-Riccati equation, iteration/algorithm.

% Appendix: the toy model. 
% Appendix: some properties of B?

\section{Discontinuous Galerkin Approximation of the Generator of a Fluid Queue}
	\label{sec:DG}
Discontinuous Galerkin (DG) methods can be used to approximate the solutions to systems of partial differential equations (PDEs). For a more thorough description of these methods see \citep{nodalDGBook}. The domain of approximation is partitioned into intervals, referred to individually as \textit{cells} and collectively as a \textit{mesh}. On each cell, we have a finite element approximation, which constructs a finite-dimensional smooth Sobolev space using piecewise-polynomial basis functions, and then projects the partial differential equations onto this space. This projection leads to a new system of equations, referred to as the \textit{weak form} of the original system of PDEs. Next, we must approximate the \textit{flux operator} which moves probability from one cell to another, in a manner similar to the underlying principle of a finite-volume approximation. This method conserves probability, and can handle discontinuities, such as jumps and point masses. Here we construct the DG approximation to the matrix of operators \(\mathbb B\) which we use later to construct a DG approximation to \(\mathbb D(s)\) then \(\mathbb\Psi(s)\), and ultimately the stationary distribution of an SFFM.

\subsection{The Partial Differential Equation}
We start by introducing the PDE from which we will extract the approximation to the generator \(\mathbb B\).

Let $f_i(x,t)$ be the joint density of $\{(X_t, \varphi_t)\}$: 
	% 
	\begin{align*} 
	    f_i(x,t) := \frac{\partial}{\partial x} 	\mathbb{P}\left(X_t \leq x, \varphi_t = i\right),\quad x>0,\, i \in\calS,
	\end{align*} 
which satisfies the system of partial differential equations 
%
\begin{align*}
\frac{\partial}{\partial t} f_i(x,t)& = \sum_{j\in \mathcal{S}}  f_j(x,t)T_{ji} - c_i \frac{\partial}{\partial x} f_i(x,t), \quad x>0, i\in\mathcal S
\end{align*}
% 
subject to suitable boundary conditions~\citep{bo2014}. In matrix form, 
\begin{align}
\cfrac{\partial}{\partial t} \boldsymbol f(x,t) &= \boldsymbol f(x,t)\bs T -  \cfrac{\partial}{\partial x}\boldsymbol f(x,t)\bs C\label{eqn:pde_density_matrix}, 
\end{align}
where \(\boldsymbol f(x,t) = \left(f_i(x,t)\right)_{i\in\mathcal S}\) is a row-vector. 
This system of PDEs is closely related to the generator \(\mathbb B\). Write \(\boldsymbol \mu(t)(\cdot) = (\mu_j(t)(\cdot))_{j\in\mathcal S}\), then, for \(\mathcal A\subset (0,\infty)\), and assuming \(\boldsymbol \mu(t)\) admits a density, 
\begin{align*}
	\boldsymbol \mu(t)(\mathcal A) = \int_{x\in\mathcal A}\boldsymbol f(x,t)\wrt x.
\end{align*}
That is, \(f_i(x,t)\) is the density of \(\mu_i(t)\). 
Then \(\boldsymbol \mu(t)\) satisfies the operator differential equation
\begin{align*}
	\cfrac{\wrt}{\wrt t}\boldsymbol \mu(t)(\wrt x) = \boldsymbol \mu(t)\mathbb B(\wrt x) = \boldsymbol f(x,t)\bs T\wrt x - \cfrac{\partial}{\partial x}\boldsymbol f(x,t)\bs C\wrt x,
\end{align*}
on the interior of the space \([0,\infty)\). Thus, by approximating the operator on the right-hand side of Equation (\ref{eqn:pde_density_matrix}) we can approximate the infinitesimal operator \(\mathbb B\). The DG method does exactly this, by approximating the operator with a matrix.

\subsection{Cells, Test Functions, and Weak Formulation}
To begin with, consider an unbounded first fluid level \(\{\widehat X_t,t\geq0\}\), \(\widehat X_t\in(-\infty,\infty)\). We will eventually truncate this space so that we have a finite dimensional approximation; however, this requires a discussion on boundary conditions which we save for later. Let \(\mathcal D_k = [x_k,x_{k+1}],\, k\in\mathbb Z\) partition the domain \((-\infty,\infty)\). We call the \(\calD_k\) \textit{cells}.

On each cell \(\calD_k\) we choose \(N_k\) linearly independent functions \(\{\phi_r^k\}_{r=1}^{N_k}\), compactly supported on \(\calD_k\) (i.e.~\(\phi_r^k(x)=0\) for \(x\notin\calD_k\)) to form a basis for the space \(W_k\), in which we formulate the approximation. Here, as is standard in DG methods \citep{nodalDGBook}, we take \(\{\phi_r^k\}_{r=1}^{N_k}\) to be the space of polynomials of degree \(N_k-1\). It is convenient in this work to take \(\{\phi_r^k\}_{r=1}^{N_k}\) as a basis of Lagrange interpolating polynomials defined by the Gauss-Lobatto quadrature points, since our approximations inherit nice properties from this \citep{nodalDGBook}. However, the constructions presented here are general and any basis can be used. For the sake of illustration, the reader may think of \(\{\phi_r^k\}_{r=1}^{N_k}\) as the Lagrange polynomials. On each cell \(\mathcal D_k\) we approximate 
\[f_i(x,t)\approx u_i^k(x,t)=\sum\limits_{r=1}^{N_k}a_{i,r}^k(t)\phi_r^k(x),\] 
where \(a_{i,r}^k(t)\) are yet-to-be-determined time-dependent coefficients. We refer to \(u_i^k\) as the \textit{local} approximation on \(\calD_k\), while the \textit{global} approximation is given by \(\sum\limits_{k\in\mathbb Z}u_i^k\) on the whole domain. The whole approximation space is \(\bigoplus\limits_{k\in\mathbb Z} W_k\).

Let \(\mathcal N_k := \left\{1,\dots,N_k\right\},\, k \in \mathbb Z\). For \(k\in\mathbb Z,\) define \textit{local} row-vectors 
\[\boldsymbol \phi^k(x) = (\phi_r^k(x))_{r\in\mathcal N_k}, \quad \boldsymbol a_i^k(x) = (a_{i,r}^k(x))_{r\in\mathcal N_k},\,i\in\mathcal S.\]
%Also define \textit{global} row-vectors 
%\[\boldsymbol \phi(x) = (\boldsymbol\phi^k(x))_{k\in\{1,...,K\}},\quad \boldsymbol a_i(x) = (\boldsymbol a_{i}^k(x))_{k\in\{1,...,K\}},\,i\in\mathcal S.\]
Note that we will always use the letter \(r\) to index the basis function within each cell.

The DG method proceeds by first considering the \textit{weak-formulation} of the PDE, which is constructed from the \textit{strong-form} of the PDE, Equation (\ref{eqn:pde_density_matrix}). In general, to construct the weak-form we need a set of test functions, say \(W\). Now, take the strong form of the PDE, multiply it by some test function \(\psi(x)\in W\), integrate with respect to \(x\), and apply integration by parts to the derivative with respect to \(x\), to get 
\begin{align}
\begin{multlined}[t]
	\int_{x\in\mathbb R}\cfrac{\partial}{\partial t} f_j(x,t)\psi(x)\wrt x = \int_{x\in\mathbb R} \sum_{i\in\calS}f_i(x,t)T_{ij}\cfrac{\wrt}{\wrt x}\psi(x)\wrt x 
	%
	+  \int_{x\in\mathbb R} f_j(x,t)c_j\psi(x)\wrt x \\{}- [ f_j(x,t)c_j\psi(x)]_{x=-\infty}^{x=\infty}, \end{multlined}\label{eqn:weak form}
\end{align}
for \(j\in\calS\). It is common to choose \(W\) such that \(\psi(-\infty)=\psi(\infty)=0\), in which case the last term on the right is zero. Requiring (\ref{eqn:weak form}) to hold for every \(\psi\in W\) gives the weak-formulation of the PDE. For a sufficiently rich set of test functions \(W\) the weak and strong forms of the PDE are equivalent. Solutions to (\ref{eqn:weak form}) are known as \textit{weak} solutions and generalise the concept of a solution of the PDE. For example, this may allow discontinuities with respect to \(x\) in the solution -- something which is ill-defined for the strong form.

For the purpose of DG, we take the set of test functions to be \(W = \bigoplus\limits_{k\in\mathbb Z} W^k\), the same as the set of basis functions of our solution space. Proceeding as described above, the weak formulation is 
\begin{align*}
	\begin{multlined}[t]\int_{x\in\calD_k}\cfrac{\partial}{\partial t} f_j(x,t)\phi_r^k(x)\wrt x = \int_{x\in\calD_k}\sum_{i\in\calS} f_i(x,t)T_{ij}\phi_r^k(x)\wrt x  
	%
	+  \int_{x\in\calD_k} f_j(x,t)c_j\cfrac{\wrt}{\wrt x}\phi_r^k(x)\wrt x \\ {}- [f_j(x,t)c_j\phi_r^k(x)]_{x=x_k}^{x=x_{k+1}}, \end{multlined}
\end{align*}
since \(\phi_r^k\) is compactly supported on \(\calD_k\), for all \(j\in\mathcal S,\,r\in\mathcal N_k\), \(k\in\mathbb Z.\) Now, note that any function \(g(x)\) can be decomposed as \(g(x) = g^{W}(x)+g^\perp(x)\) where \(g^{W}\in W\) and \(g^\perp \in W^\perp\), and \(W^\perp\) is the orthogonal complement of \(W\). Since \(g^\perp\) is orthogonal to \(W\), \(\displaystyle\int_{x}g^\perp(x)\phi_r^k(x)\wrt x=0\) for \(r\in\mathcal N_k,\,k\in\mathbb Z\). Also, note that \(\cfrac{\wrt}{\wrt x}\phi_r^k(x)\in W\). Using this, we can write 
\begin{align*}
	\begin{multlined}[t]\int_{x\in\calD_k}\cfrac{\partial}{\partial t} \left(f_j^W(x,t)+f_j^\perp(x,t)\right)\phi_r^k(x)\wrt x 
	= \int_{x\in\calD_k}\sum_{i\in\calS} \left(f_i^W(x,t)+f_i^\perp(x,t)\right)T_{ij}\phi_r^k(x)\wrt x  
	%
	\\ {}
	+  \int_{x\in\calD_k} \left(f_j^W(x,t)+f_j^\perp(x,t)\right)c_j\cfrac{\wrt}{\wrt x}\phi_r^k(x)\wrt x - [f_j(x,t)c_j\phi_r^k(x)]_{x=x_k}^{x=x_{k+1}}, \end{multlined}
\end{align*}
	which is equivalent to
\begin{align}
	\begin{multlined}[t]\int_{x\in\calD_k}\cfrac{\partial}{\partial t} f_j^W(x,t)\phi_r^k(x)\wrt x = \int_{x\in\calD_k}\sum_{i\in\calS} f_i^W(x,t)T_{ij}\phi_r^k(x)\wrt x  
	%
	+  \int_{x\in\calD_k} f_j^W(x,t)c_j\cfrac{\wrt}{\wrt x}\phi_r^k(x)\wrt x \\{} - [f_j(x,t)c_j\phi_r^k(x)]_{x=x_k}^{x=x_{k+1}}. \end{multlined} \label{eqn:hash}
\end{align}
Now, \(f_j^W(x,t)\in W\) so, on \(\mathcal D_k\), it can be expressed as \(\boldsymbol a_j^k(t) \boldsymbol \phi^k(x)\tr{}\), which we now substitute into (\ref{eqn:hash})
%\begin{align*}
%	\begin{multlined}[t]\int_{x\in\calD_k}\cfrac{\wrt}{\wrt t} \boldsymbol a_j^k(t) \boldsymbol \phi^k(x)\tr{}\phi_r^k(x)\wrt x = \int_{x\in\calD_k}\sum_{i\in\calS} \boldsymbol a_i^k(t) \boldsymbol \phi^k(x)\tr{}T_{ij}\phi_r^k(x)\wrt x  
%	%
%	\\ {}+  \int_{x\in\calD_k}\boldsymbol a_j^k(t) \boldsymbol \phi^k(x)\tr{}c_j\cfrac{\wrt}{\wrt x}\phi_r^k(x)\wrt x - c_j[f_j(x,t)\phi_r^k(x)]_{x=x_k}^{x=x_{k+1}}.\end{multlined}
%\end{align*}
and repeat this for all test functions \(\phi_r^k(x)\), \(r=1,...,N_k\), to get the following system of equations,
\begin{align}
	\begin{multlined}[t]\int_{x\in\calD_k}\cfrac{\wrt}{\wrt t} \boldsymbol a_j^k(t) \boldsymbol \phi^k(x)\tr{}\boldsymbol \phi^k(x)\wrt x = \int_{x\in\calD_k}\sum_{i\in\calS} \boldsymbol a_i^k(t) \boldsymbol \phi^k(x)\tr{}T_{ij}\boldsymbol\phi^k (x)\wrt x  
	%
	\\{}+  \int_{x\in\calD_k}\boldsymbol a_j^k(t) \boldsymbol \phi^k(x)\tr{}c_j\cfrac{\wrt}{\wrt x}\boldsymbol\phi^k(x)\wrt x - c_j[f_j(x,t) \boldsymbol\phi^k(x)]_{x=x_k}^{x=x_{k+1}},\quad k\in\mathbb Z. \end{multlined} \label{eqn:DG system}
\end{align}

\subsection{Mass, Stiffness, and Flux}
For \(k\in\mathbb Z\), define local \textit{mass} and \textit{stiffness} matrices \(\bs M_k\) and \(\bs G_k\) by 
\[\bs M_k := \int_{x\in\calD_k}\boldsymbol \phi^k(x)\tr{}\boldsymbol \phi^k(x)\wrt x,\quad \bs G_k := \int_{x\in\calD_k}\boldsymbol \phi^k(x)\tr{}\cfrac{\wrt}{\wrt x} \boldsymbol \phi^k(x)\wrt x.\]
%and global mass and stiffness matrices by
%\begin{align*}
%	M &:= \left[\begin{array}{ccc} M_1 & & \\ & \ddots & \\ & & M_K \end{array}\right] = \int_{x\in\mathcal I}\boldsymbol \phi(x)\tr{}\boldsymbol \phi(x)\wrt x,
%	\\G &:= \left[\begin{array}{ccc} G_1 & & \\ & \ddots & \\ & & G_K \end{array}\right] = \int_{x\in\mathcal I}\boldsymbol \phi(x)\tr{}\cfrac{\wrt}{\wrt x}\boldsymbol \phi(x)\wrt x,
%\end{align*}
We can write (\ref{eqn:DG system}) as 
\begin{align}
	&\cfrac{\wrt}{\wrt t} \boldsymbol a_j^k(t) \bs M_k = \sum_{i\in\calS} \boldsymbol a_i^k(t)M_kT_{ij} 
	%
	+  c_j\boldsymbol a_j^k(t) \bs G_k - c_j[f_j(x,t)\boldsymbol\phi^k(x)]_{x=x_k}^{x=x_{k+1}}.\label{eqn:DG matrix}
\end{align}

It remains to approximate the \textit{flux}, \(f_j(x,t)\) at the cell edges \(x_k,\,k\in\mathbb Z\), so that we may evaluate the terms \([f_j(x,t)\phi_r^k(x)]_{x=x_k}^{x=x_{k+1}}\), \(r=1,...,N_k,\,k\in\mathbb Z\). This is the key for DG -- it joins the local approximations on each cell \(\calD_k\), into a global approximation on the whole domain of approximation. The flux is the instantaneous rate (with respect to time) at which density moves across the boundaries \(x_k,\,k\in\mathbb Z\). There are different choices for the flux, and we refer the reader to \citep{c99,nodalDGBook}, and references therein, for some discussion of the topic. Here, we choose the \textit{upwind} scheme, which, as we shall see, closely resembles the flux terms from the generator \(\mathbb B\). The approximate flux, also known as the \textit{numerical flux}, is given by 
\[f^*_j(x,t) = sign(c_j)\lim_{\varepsilon\to0^+}u_j(x-\varepsilon c_j,t),\]
at each \(x=x_k,\,k\in\mathbb Z\). 
Intuitively, the upwind flux takes the value of the density immediately on the upwind side of each \(x_k\).

Denote by \(x^-\) and \(x^+\) the left and right limits at \(x\), respectively. Assume first \(c_j>0\), then 
\begin{align*}
	-c_j[f_j(x,t)\phi_r^k(x)]_{x=x_k}^{x=x_{k+1}} & \approx -c_j[f_j^*(x,t)\phi_r^k(x)]_{x=x_k}^{x=x_{k+1}}
	\\& = -c_jf_j^*(x_{k+1},t)\phi_r^k(x_{k+1}) + c_jf_j^*(x_{k},t)\phi_r^k(x_{k})
	\\& = -c_ju_j(x_{k+1}^-,t)\phi_r^k(x_{k+1}) + c_ju_j(x_{k}^-,t)\phi_r^k(x_{k})
	\\& = -c_ju_j^k(x_{k+1}^-,t)\phi_r^k(x_{k+1}) + c_ju_j^{k-1}(x_{k}^-,t)\phi_r^k(x_{k})
	\\& = -c_j\boldsymbol a_j^k(t)\boldsymbol \phi^k(x_{k+1}^-)\tr{}\phi_r^k(x_{k+1}) + c_j\boldsymbol a_j^{k-1}(t)\boldsymbol \phi^{k-1}(x_{k}^-)\tr{}\phi_r^k(x_{k}).
\end{align*}
In matrix form,  
\begin{align*}
	-c_j[f_j(x,t)\boldsymbol\phi^k(x)]_{x=x_k}^{x=x_{k+1}} & \approx -c_j[f_j^*(x,t)\boldsymbol\phi^k(x)]_{x=x_k}^{x=x_{k+1}}
	\\& = -c_j\boldsymbol a_j^k(t)\boldsymbol \phi^k(x_{k+1}^-)\tr{}\boldsymbol\phi^k(x_{k+1}) + c_j\boldsymbol a_j^{k-1}(t)\boldsymbol \phi^{k-1}(x_{k}^-)\tr{}\boldsymbol\phi^k(x_{k})
	\\& = c_j\boldsymbol a_j^k(t)F_j^{k,k}+c_j\boldsymbol a_j^{k-1}(t)F_j^{k-1,k},
\end{align*}
where, for \(j\in \calS\) with \(c_j>0\), we define \(\bs F_j^{k,k} := -\boldsymbol \phi^k(x_{k+1}^-)\tr{}\boldsymbol\phi^k(x_{k+1}),\,k\in\mathbb Z\) and \(\bs F_j^{k-1,k} := \boldsymbol \phi^{k-1}(x_{k}^-)\tr{}\boldsymbol\phi^k(x_{k}),\,k\in\mathbb Z\).

Now proceed similarly for \(c_j<0\) to get the approximation 
\begin{align*}
	-c_j[f_j(x,t)\boldsymbol\phi^k(x)]_{x=x_k}^{x=x_{k+1}} & \approx -c_j[f_j^*(x,t)\boldsymbol\phi^k(x)]_{x=x_k}^{x=x_{k+1}}
	\\& = -c_j\boldsymbol a_j^{k+1}(t)\boldsymbol \phi^{k+1}(x_{k+1}^+)\tr{}\boldsymbol\phi^k(x_{k+1}) + c_j\boldsymbol a_j^{k}(t)\boldsymbol \phi^{k}(x_{k}^+)\tr{}\boldsymbol\phi^k(x_{k})
	\\& = c_j\boldsymbol a_j^{k+1}(t)\bs F_j^{k+1,k}+c_j\boldsymbol a_j^{k}(t)\bs F_j^{k,k},
\end{align*}
where, for \(j\in \calS\) with \(c_j<0\), we define \(\bs F_j^{k+1,k} := -\boldsymbol \phi^{k+1}(x_{k+1}^+)\tr{}\boldsymbol\phi^k(x_{k+1}),\,k\in\mathbb Z,\) and \(\bs F_j^{k,k} := \boldsymbol \phi^{k}(x_{k}^+)\tr{}\boldsymbol\phi^k(x_{k}),\,k\in\mathbb Z\).

The matrices \(\bs F_j^{k-1,k},\,\bs F_j^{k,k},\) and \(\bs F_j^{k+1,k}\) are the local flux matrices. For convenience, we also define the matrices \(\bs F_j^{k,k+1}=0\) for \(c_j<0\) and \(\bs F_j^{k,k-1}=0\) for \(c_j>0\), \(k\in\mathbb Z\).

% from which we can assemble the global flux matrices
%\begin{align*}
%	F_j &= \left[\begin{array}{cccccc}
%		F_j^{1,1} & F_j^{1,2} & & & & \\
%		 & F_j^{2,2} & F_j^{2,3} & & & \\
%		& & \ddots & \ddots & & \\
%		& & & \ddots & \ddots & \\
%		& & & & F_j^{K-1,K-1} & F_j^{K-1,K} \\
%		& & & & & F_j^{K,K}
%	\end{array}\right],\quad c_j\geq0,
%	%
%	\\F_j &= \left[\begin{array}{cccccc}
%		F_j^{1,1} & & & & & \\
%		F_j^{2,1} & F_j^{2,2} & & & & \\
%		& \ddots & \ddots & & & \\
%		& & \ddots & \ddots & & \\
%		& & & F_j^{K-1,K-2} & F_j^{K-1,K} & \\
%		& & & & F_j^{K-1,K} & F_j^{K,K}
%	\end{array}\right],\quad c_j<0.
%\end{align*}
%
%We can now write 
%\begin{align*}
%	&\cfrac{\wrt}{\wrt t} \boldsymbol a_j(t) M = \sum_{i\in\calS} \boldsymbol a_j(t)MT_{ij} 
%	%
%	+  c_j\boldsymbol a_j(t) (G + F_j).
%\end{align*}
%
%Defining \(\boldsymbol a(t) = (\boldsymbol a_j(t))_{i\in\calS}\), then 
%\begin{align}\label{eqn:DG ODE}
%	&\cfrac{\wrt}{\wrt t} \boldsymbol a(t)(I_{N_\calS}\otimes M) = \boldsymbol a(t){  B,}
%\end{align}
%where 
%\[{  B} = \left[(T\otimes M) +
%	\left[\begin{array}{ccc}
%		c_1(G+F_1) & &  \\
%		& \ddots & \\
%		& & c_{N_\calS}(G+F_{N_\calS})\\
%	\end{array}\right] \right],\]
%\(\otimes\) is the kronecker product and \(I_{N_\calS}\) is the \(N_\calS\times N_\calS\) identity matrix.

%So far, this DG approximation has been constructed according to the lexicographic ordering of \(\calS\times\{1,...,K\}\). A reordering of this construction according to the lexicographic ordering of \(\{1,...,K\}\times\calS\) helps elucidate the connection with the partitioned operator \(\mathbb B\) and eases notation slightly for the next discussion on boundary conditions.

%Also define \textit{global} row-vectors 
%\[\boldsymbol \phi(x) = (\boldsymbol\phi^k(x))_{k\in\{1,...,K\}},\quad \boldsymbol a_j(x) = (\boldsymbol a_{i}^k(x))_{k\in\{1,...,K\}},\,i\in\mathcal S.\]

To write this out as a \textit{global} system, define the row-vectors 
\[\boldsymbol a^k(t) = (\boldsymbol a_{i}^k(t))_{i\in\mathcal S},\quad \boldsymbol a(t) = (\boldsymbol a^k(t))_{k\in\mathbb Z},\]
and the block-diagonal matrix 
\begin{align*}
\widetilde{\bs M} &= \left[\begin{array}{ccc}\ddots&&\\&\bs I_{N_\calS}\otimes \bs M_k&\\&&\ddots\end{array}\right],
\intertext{where \(N_\calS=|\calS|\), \(\otimes\) is the Kronecker product, and the block-tridiagonal matrix}
\widetilde{\mathfrak B} &= \left[\begin{array}{ccccc}
	\ddots & \ddots & \ddots & & \\
	& \widetilde{\mathfrak B}^{k,k-1} & \widetilde{\mathfrak B}^{k,k} & \widetilde{\mathfrak B}^{k,k+1} & \\
	& & \ddots & \ddots & \ddots 
\end{array}\right],
\end{align*}
where, for \(k\in\mathbb Z\), 
\begin{align*}
    \widetilde{\mathfrak{\bs B}}^{kk}&=\left[\begin{array}{ccc}T_{11}\bs M_k + c_1(\bs F_1^{kk}+\bs G_k) & T_{12}{\bs M_k} & T_{1N_\calS}{\bs M_k}   \\ T_{21}\bs M_k & & \\ \vdots &\ddots & \vdots \\ & &   T_{N_\calS-1,N_\calS}{\bs M_k} \\  T_{N_\calS1}{\bs M_k} &  T_{N_\calS,N_\calS-1}{\bs M_k} & T_{N_\calS,N_\calS}{\bs M_k} +c_{N_\calS}(\bs F_{N_\calS}^{kk}+\bs G_k)\end{array}\right],\\
	\widetilde{\mathfrak{\bs B}}^{k,k+1}&=\left[\begin{array}{ccc}c_1\bs F_1^{k,k+1}&  & \\  &\ddots & \\  &  &c_{N_\calS}\bs F_{N_\calS}^{k,k+1} \end{array}\right],\\
	\widetilde{\mathfrak{\bs B}}^{k,k-1}&=\left[\begin{array}{ccc}c_1\bs F_1^{k,k-1}&  & \\  &\ddots &  \\  &  &c_{N_\calS}\bs F_{N_\calS}^{k,k-1} \end{array}\right].
\end{align*} 
The matrices \(\widetilde{\mathfrak{\bs B}}^{k\ell}\) are defined by sub-blocks; denote these sub-blocks by \(\widetilde{\mathfrak{\bs B}}^{k\ell}_{ij}\):
\begin{align*}
	\widetilde{\mathfrak{\bs B}}^{kk}_{ij} &= \begin{cases}T_{ij}\bs M_k + c_i(\bs F_i^{kk}+\bs G_k) & i=j,\\T_{ij}\bs M_k & i\neq j,\end{cases}
\\	\widetilde{\mathfrak{\bs B}}^{k\ell}_{ij} &= \begin{cases}c_i\bs F_i^{k\ell} & i=j,\\0 & i\neq j,\end{cases}\quad \ell \in \{k-1,k\}.
\end{align*}
The global system of equations is 
\begin{align}\label{eqn:DG ODE}
	&\cfrac{\wrt}{\wrt t} \boldsymbol a(t)\widetilde{\bs M} = \boldsymbol a(t){\widetilde{\mathfrak{\bs B}}}.
\end{align}

\subsection{Boundary conditions}
To enable computation, this numerical approximation has to take place on a finite interval, which means we must consider a bounded domain and specify boundary conditions. Recall from Section~\ref{sec:prelim} that we wish to impose a regulated boundary at \(x=0\). To apply the DG method, we must truncate the state space of the first fluid at some finite interval upper bound, \([0,\mathcal I]\), for some \(\mathcal I<\infty\), and specify the boundary behaviour at \(x=\mathcal I\). Here we consider \(\mathcal I\) to be a regulated boundary. Let us denote the doubly-bounded fluid level by \(\overline X_t\). Ultimately, we wish to approximate a fluid-fluid queue where the first fluid level, \(X_t\), is bounded below at 0, only. Thus, the first step in the approximation scheme is to approximate \(X_t\) by \(\overline X_t\). The truncation of \(X_t\) to \(\overline X_t\) will result in an artificial point mass at the upper bound, which we have to address properly. It is important to choose an \(\mathcal I\) sufficiently large to control the error induced by the artificial upper bound, however, with larger \(\mathcal I\) there comes increased computational burden. We shall further comment on this in Section~\ref{sec:numexp}, where we report our numerical experiments.

Let $[0,\mathcal{I}]$ be the domain of the approximation, where $\mathcal{I} < \infty$, and assume there is a regulated boundary for \(\{\overline X_t\}\) at \(x=\mathcal I\). We partition the space $[0,\mathcal{I}]$ into \(\mathcal D_\nabla=\{0\},\) \(\mathcal D_\Delta=\{\mathcal I\},\) and \(K\) non-trivial intervals, \(\calD_k=[x_k,x_{k+1}]\setminus \{\{0\}\cup\{\mathcal I\}\},\, x_k<x_{k+1},\, k=1,...,K\), \(x_1=0,\,x_{K+1}=\mathcal I\) and define \(h_k := x_{k+1}-x_k\). The notation \(\Delta\) refers to quantities and sets which are relevant to the boundary at \(\mathcal I\).

For states with \(c_i\leq 0\), there is the possibility of point mass accumulating at the boundary at~\(0\). Denote these point masses by \(q_{\nabla,i}(t)\) for \(i\in\mathcal S_\nabla\). For states with \(c_i>0\) there is no possibility of a point mass at \(0\). Similarly, for \(c_i\geq 0\) there is the possibility of a point mass at \(\mathcal I\). Denote these point masses by \(q_{\Delta,i}(t)\), for \(i\in\mathcal S_\Delta\). For states with \(c_i<0\) there is no possibility of a point mass at \(\mathcal I\). Let \(\bs q_\nabla(t)=(q_{\nabla,i}(t))_{i\in\calS_\nabla}\) and \(\bs q_\Delta(t) = (q_{\Delta,i}(t))_{i\in\calS_\Delta}\) and \(\bs f_m(x,t) = (f_i(x,t))_{i\in\calS_m}\), \(m\in\{+,-,0\}\).

Let us first consider the boundary at \(\overline X_t=0\). \cite{bo2014} show the following boundary conditions describe the evolution of probability/density of a stochastic fluid model with a regulated boundary at \(0\);
\begin{align}\label{eqn:BC1}
\cfrac{\wrt}{\wrt t}\bs q_\nabla(t) &= \bs q_\nabla(t) T_{\nabla \nabla} - \boldsymbol f_\nabla(0,t)C_\nabla,
\\\label{eqn:BC2}
\bs q_\nabla(t)T_{\nabla+}&=\bs f_+(0,t)C_+.
\end{align}
Equation (\ref{eqn:BC1}) states that point mass moves between phases according to the sub-generator matrix \(T_{\nabla\nabla}\), and that the flux of probability density into the point masses is \(- \boldsymbol f_\nabla(0,t)C_\nabla\). Substituting the DG approximation for \(\boldsymbol f_\nabla(0,t)\) into (\ref{eqn:BC1}) gives, for \(j\in\calS_\nabla\), 
\begin{equation*}\label{eqn:DGBC1}
\cfrac{\wrt}{\wrt t} q_{\nabla,j}(t) = \sum_{i\in\calS_\nabla} q_{\nabla,i}(t) T_{i j} - \boldsymbol a^1_j(t)\boldsymbol \phi^1(0)\tr{}c_j.
\end{equation*}
Equation (\ref{eqn:BC2}) describes the flux of probability mass to density upon a transition from a phase in \(\calS_\nabla\) to a phase in \(\calS_+\). Thus, the flux into the left-hand edge of \(\calD_1\) in phase \(j\in\calS_+\) is, \(\sum\limits_{i\in\calS_\nabla} q_{\nabla,i}(t)T_{ij}\). Therefore, we can now evaluate 
\begin{align*}
	-c_j[f_j(x,t)\boldsymbol\phi^1(x)]_{x=0}^{x=x_{1}} & =  -c_jf_j(x_1,t)\boldsymbol\phi^1(x_1)+c_jf_j(0,t)\boldsymbol\phi^1(0)
	\\& \approx -c_j(f_j^*(x_1,t)\boldsymbol\phi^1(x_1)+\sum_{i\in\calS_\nabla}q_{\nabla,i}(t)T_{ij}\boldsymbol\phi^1(0)
	\\& = c_j\boldsymbol a_j^{1}(t)F_j^{1,1}+\sum_{i\in\calS_\nabla}q_{\nabla,i}(t)T_{ij}\boldsymbol\phi^1(0),
\end{align*}
for \(i \in \calS_+\). 
Thus, the DG approximation of the flux into point masses \(q_{\nabla,j}(t)\) is \(-\boldsymbol a_j^1(t)\boldsymbol \phi^1(0)\tr{} c_j,\,j\in\calS_-\), the rate of transition of point mass within \(\bs q_{\nabla}(t)\) is \(T_{\nabla\nabla}\), and the DG approximation of the transition of point mass to density is \(\sum\limits_{i\in\calS_\nabla}q_{\nabla,i}(t)T_{ij}\boldsymbol\phi^1(0),\,j\in\calS_+\).

Similarly, for the boundary at \(X_t=\mathcal I\) the boundary conditions are 
\begin{align*}
\cfrac{\wrt}{\wrt t}\bs q_\Delta(t) &= \bs q_\Delta(t) T_{\Delta \Delta} + \boldsymbol f_\Delta(\mathcal I,t)C_\Delta,\\
\bs q_\Delta(t)T_{\Delta-}&=-\bs f_-(\mathcal I,t)C_-.
\end{align*}
Using the same arguments as above, 
\begin{align*}
\cfrac{\wrt}{\wrt t} q_{\Delta,j}(t) &= \sum_{i\in\calS_\Delta}q_{\Delta,i}(t) T_{ij} + \boldsymbol a^K_j(t)\boldsymbol \phi^K(\mathcal I)\tr{}c_j,
\\-c_j[f_j(x,t)\boldsymbol\phi^K(x)]_{x=x_K}^{x=\mathcal I} & \approx c_j\boldsymbol a_j^{K}(t)F_j^{K,K}+\sum_{i\in\calS_\Delta}q_{\Delta,i}(t)T_{ij}\boldsymbol\phi^K(\mathcal I),
\end{align*}
for \(j\in\calS_\Delta\). 
Thus, the DG approximation of the flux into the point mass \(q_{\Delta,j}(t)\) is \(\boldsymbol a_j^K(t)\boldsymbol \phi^K(0)\tr{} c_j\), \(j\in\calS_+\), the rate of transition of point mass within \(\bs q_{\Delta}(t)\) is \(T_{\Delta\Delta}\), and the DG approximation of the transition of point mass to density is \(\sum\limits_{i\in\calS_\Delta}q_{\Delta,i}(t)T_{ij}\boldsymbol\phi^K(\mathcal I)\), \(j\in\calS_-\).
